---
title: "Exploring the factors influencing the perceived ideal timing of getting married in Sweden, 2018"
author: "Anastasia Alekhina, Ekaterina Titova, Tatiana Ivanova"
output:
  github_document:
    toc: true
    toc_depth: 3
always_allow_html: yes
date: "Last edited `r format(Sys.time(), '%d %B %Y')`"
---

```{=html}
<style>
h1.title {
color: #261D4C;
font-family: Verdana;
font-size: 36px;
}

h2 {
color: #261D4C;
font-family: Verdana;
font-size: 23px;
}

h3 {
color: #261D4C;
font-family: Verdana;
font-size: 18px;
}

#TOC {
color: #261D4C;
font-family: Verdana;
font-size: 16px;
border-color: #a897f9;
}
body {
color: #261D4C;
font-family: Verdana;
background-color: #FFFFFF;
}
pre {
color: #261D4C;
font-family: Verdana;
background-color: #FAF9FF;
border-color: #a897f9;
}
</style>
```
```{r setup, include=FALSE,  warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(ggplot2)
library(sjPlot)
library(dplyr)
library(kableExtra)
library(vcd)
library(car)
library(ggstatsplot)
library(plyr)
library(scales)
library(psych)
library(magrittr)
library(knitr)
library(kableExtra)
library(corrplot)
library(rstatix)
library(effsize)
library(DescTools) 
library(ggpubr)
library(sjstats)
webshot::install_phantomjs()

ess9 <- 
  read.spss("~/Desktop/hse/2nd YEAR/DATA ANALYSIS/data/ESS9SE.sav", 
            use.value.labels = T, 
            to.data.frame = T)
ktn<-subset(ess9, select = c("eisced", "iagmr", "iagpnt",  "iaglptn", "hinctnta", "eduyrs", "rlgdgr", "yrbrn", "gndr", "admge"))

ktn <- ktn[ -c(208, 242, 1420), ]
Mode <- function(x) {
ux <- unique(x)
ux[which.max(tabulate(match(x, ux)))]
}
```

# Topic

Exploring the impact of attitudes, sociodemographic and socio-economic factors on perceived ideal timing of getting married in Sweden

# Research question

1.  How are the household income, age, religion, attitudes toward ideal cohabitation and parenting timing, education of a person associated with the age one perceives as ideal to get married?

# Theoretical background & Research Hypotheses

### Life course and history theories

To explain how and why organisms, including humans, allocate energy, time, and resources across their lifetime, evolutionary biologists and behavioral ecologists developed **life history theory**. Energy devoted to **somatic effort** (the growth and maintenance of one's body and mind, including the accumulation of knowledge, skills, and other types of embodied capital) cannot simultaneously be devoted to **reproductive effort** (intrasexual competition, courtship, gestation, birth, and childcare) (Griskevicius et. al 2011) The particular nature of how and when this tradeoff is resolved by a given organism constitutes that organism's life history strategy.

Existing literature on marital timing examines the influence of various factors on the formation of the ideal age for marriage. Several studies have shown that **education and income levels have a positive effect on perceived ideal age to marry** and subsequently on becoming parents: people with higher levels of SES marry and start families later. This effect can be explained by **developmental life course theories**, namely that young adults in their "emerging adulthood" (age 20 and older), who are involved in educational and career processes, are less interested in marriage and family life (Fuller, Frost, and Burr 2015).

## Evolutionary theory and theory of parental investment

According to **evolutionary theory** men and women develop different gender-specific adaptive mechanisms, resulting from fact of facing different mating and sexual behavior problems. Timing of adulthood/maturity and reproduction-related behavior emerge as major evolutionary mechanisms based on gender differences (Buss & Schmitt, 2011). Women and men thus tend to determine differently the optimal maturity age for both their own and the opposite sex.

Trivers's **theory of parental investment** explains some possible mechanisms by which this determination of age might operate. According to the theory, for gender with greater parental investment the **resources of the opposite sex prove to be a powerful factor**, in consequence of which women will tend to set a higher maturity and reproductive-related age for men than men do for themselves. (Trivers, 1972; Pawlowski & Dunbar, 1999; Buss & Schmitt, 1993) Conversely, men, being on the lower parental investment side, will be motivated by their desire to date younger women because of their physical characteristics and attractiveness as indicators of their reproductive value for instance. Whereas women themselves will be inclined to indicate a lower age for themselves, since they believe that it takes time to find a worthy partner who can ensure the family.

### Parenting and marriage

Cherlin and colleagues created a study where they test three propositions regarding mothers' attitudes toward childbearing, marriage, and divorce (Cherlin et. al, 2008). They suggested that **mothers' attitudes and preferences in these three domains do not support the long delay between childbearing and marriage** that has been noted in the literature. Indeed, they found that 59 percent of the sample chose the same 5-year age category for each of the 2 questions ("Best Age to Marry" and "Best Age to Start Having Children"), which implies that a majority of the mothers thinks the **best ages to start childbearing and to marry are within 5 years of each other**. According to this article, we can justify our proposition to test the influence of an ideal age for getting pregnant.

### Cohabitation and marriage

Willoughby and Carroll in their study examined whether attitudes toward cohabitation would be related to attitudes toward marriage (Willoughby and Carroll, 2011). It was found that a later ideal age of **marriage was associated with a stronger belief that cohabitation was beneficial**. In addition, young adults who chose a younger ideal age of marriage indicated that marriage was an important personal goal were less likely to approve of cohabitation without marriage plans. Young adults who value marriage and want to get married early may view cohabitation as an **acceptable step toward a marriage relationship**, but may be less likely to approve of cohabitation without marriage, and see it as an alternative to being single or practicing serial cohabitation. Based on this article, we can test whether the ideal age of starting cohabitation with a partner will affect the ideal age of marriage, as people view cohabitation as one of the important stages in their lives before marriage.

### Religiosity

Fuller and Frost found the pervasive effect of religiosity on perceived marital timing, namely their results show a **negative association between religiosity with perceived ideal age to marry and ideal time to marry**. Thus, they found that with increasing levels of religiosity, respondents were most likely to report an earlier ideal age to marry and an earlier ideal time to marry (Fuller, Frost, and Burr 2015).

### Trend toward desiring later marriage

According to Thornton and Young-DeMarco (2001), the trend toward desiring later marriage continued from the mid-1970s through the 1990s. Therefore, people who participated in the 70s in this research, now they are older but with the belief of early marriage as desirable one. Moreover, people who participated in the 90s in this research, they are younger now than the previously mentioned people, but respondents from the 90s share the idea of later marriage. Hence, we can test this claim by looking at the age of respondents and their perception of ideal age for getting married.

# Central measures analysis

#### Hypothesis 1

We hypothesize that mean, median and mode of perceived ideal age to get married and to become a parent will be lower for women than for men from evaluations by both genders. We expect to see a left-skewed distribution in men's and right-skewed one in women's case.

#### Hypothesis 2

We hypothesize that mean, median and mode of men perceptions of ideal age to get married and to become a parent for women will be lower than that of women themselves.

#### Hypothesis 3

We hypothesize that mean, median and mode of women perceptions of ideal age to get married and to become a parent for men will be higher than that of men themselves.

#### Hypothesis 4

We hypothesize that people with higher educational level will choose later age for marriage and becoming a parent more frequently.

#### Hypothesis 5

We hypothesize the correlation of income and the indicated ideal marriage and becoming a parent timing will be characterized by a positively sloping line.

## Variables

Prior to forming all the necessary categories and manipulating variables, four lines were removed from the data, the values of which on the main issues for us (the ideal age for marriage and the ideal age for becoming a parent) were either missing or were irrelevant within the family formation (such values as 2 for marriage, or 78 for having a child).

```{r warning = FALSE, message=FALSE}
ktn$iagmr <- as.numeric(as.character(ktn$iagmr))
ktn$iagpnt <- as.numeric(as.character(ktn$iagpnt))

# Ideal age for marriage
ktn$iagmr_cat[ktn$iagmr <="18"] <- "18 and earlier"
ktn$iagmr_cat[ktn$iagmr >= "19"& ktn$iagmr <="25"] <- "from 19 to 25"
ktn$iagmr_cat[ktn$iagmr >= "26"& ktn$iagmr <"30"] <- "from 26 to 30"
ktn$iagmr_cat[ktn$iagmr >= "31"] <- "over 31"

# Ideal age for becoming a parent 
ktn$iagpnt_cat[ktn$iagpnt <="18"] <- "18 and earlier"
ktn$iagpnt_cat[ktn$iagpnt >= "19"& ktn$iagpnt <="25"] <- "from 19 to 25"
ktn$iagpnt_cat[ktn$iagpnt >= "26"& ktn$iagpnt <"30"] <- "from 26 to 30"
ktn$iagpnt_cat[ktn$iagpnt >= "31"] <- "over 31"

```

Consequently, comparison of the reported ideal age of family formation occurred according to the 4 main criterion of respondents:

### Gende


```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x =gndr)) + geom_bar(fill = "#6F7CDF")+
  labs(x="Gender", y="Number of people")

```

The gender subgroups are evenly distributed into two categories, amounting to 778 for men and 757 for women, respectively.


### Subgroup on the question (response about females or males)

```{r warning = FALSE, message=FALSE}
#Gender subgroups datasets
f1 <- dplyr::filter(ktn, ktn$admge == "Group 1" & ktn$gndr == "Female")
m1 <- dplyr::filter(ktn, ktn$admge == "Group 1" & ktn$gndr == "Male")
f2 <- dplyr::filter(ktn, ktn$admge == "Group 2" & ktn$gndr == "Female")
m2 <- dplyr::filter(ktn, ktn$admge == "Group 2" & ktn$gndr == "Male")

f <- dplyr::filter(ktn, ktn$admge == "Group 1")
m <- dplyr::filter(ktn, ktn$admge == "Group 2")

#Gender subgroups 
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Female"] <- "Women about Women"
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Male"] <- "Men about Women"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Female"] <- "Women about Men"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Male"] <- "Men about Men"

ktn$admge_cat2[ktn$admge == "Group 1"] <- "About Women"
ktn$admge_cat2[ktn$admge == "Group 2"] <- "About Men"


ggplot(ktn, aes(x = admge_cat)) + geom_bar(fill = "#6F7CDF") + coord_flip() + scale_x_discrete(limits = rev(levels(ktn$admge_cat)))+
labs(x = "Categories", y = "Number of people")
```

By considering which answer group the person is in (group 1 or group 2) and comparing this to the gender of the person him/herself, we generated 4 subgroups, likewise evenly distributed, comprising about 380-400 people, with the category of women about men being the lowest with 357 people

### Income

```{r warning = FALSE, message=FALSE}
# Income decile
ktn$hinctnta_cat[ktn$hinctnta == "J - 1st decile"] <- "1"
ktn$hinctnta_cat[ktn$hinctnta == "R - 2nd decile"] <- "2"
ktn$hinctnta_cat[ktn$hinctnta == "C - 3rd decile"] <- "3"
ktn$hinctnta_cat[ktn$hinctnta == "M - 4th decile"] <- "4"
ktn$hinctnta_cat[ktn$hinctnta == "F - 5th decile"] <- "5"
ktn$hinctnta_cat[ktn$hinctnta == "S - 6th decile"] <- "6"
ktn$hinctnta_cat[ktn$hinctnta == "K - 7th decile"] <- "7"
ktn$hinctnta_cat[ktn$hinctnta == "P - 8th decile"] <- "8"
ktn$hinctnta_cat[ktn$hinctnta == "D - 9th decile"] <- "9"
ktn$hinctnta_cat[ktn$hinctnta == "H - 10th decile"] <- "10"

ktn$hinctnta_cat_sorted <- factor(ktn$hinctnta_cat, ordered = TRUE,
levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))

ktn$hinctnta_cat <- as.numeric(ktn$hinctnta_cat)

ggplot(na.omit(ktn), aes(x = hinctnta_cat_sorted)) + geom_bar(fill = "#6F7CDF") +
labs(x = "Income decile", y = "Number of people")

```

Respondents' incomes are compiled using the decile income distribution from the original database.  Individuals are arranged in ascending order of their income and divided into ten groups of equal size. Each income decile contains 10 percent of the population; the higher the decile, the higher the income level.



### Education

```{r warning = FALSE, message=FALSE}
# education levels
ktn$eisced_cat[ktn$eisced == "ES-ISCED I , less than lower secondary"|
                 ktn$eisced == "ES-ISCED II, lower secondary"] <- "low education"
ktn$eisced_cat[ktn$eisced =="ES-ISCED IIIa, upper tier upper secondary" |
                 ktn$eisced =="ES-ISCED IIIb, lower tier upper secondary"|
                 ktn$eisced =="ES-ISCED IV, advanced vocational, sub-degree"] <- "middle education"
ktn$eisced_cat[ktn$eisced == "ES-ISCED V1, lower tertiary education, BA level" |
                 ktn$eisced == "ES-ISCED V2, higher tertiary education, >= MA level"] <- "high education"

ktn$eisced_cat_sorted <- factor(ktn$eisced_cat, ordered = TRUE,
levels = c("low education", "middle education", "high education"))

ggplot(na.omit(ktn), aes(x =eisced_cat_sorted)) + geom_bar(fill = "#6F7CDF")+
  labs(x="Level of education", y = "Number of people")
  


```

The level of education has been distributed according to International Standard Classification of Education, where ISCED levels 0-2 correspond to low education, ISCED levels 3-4 correspond to middle education, ISCED levels 5-8 correspond to higher education

Our data ended up having 1535 observations and 14 variables, 6 of which were taken directly from the ESS9 database, while the other 8 were created.
``` {r warning = FALSE, message=FALSE}
cls1 <- class(ktn$hinctnta_cat_sorted)
cls2 <- class(ktn$eisced_cat_sorted)

Table <- matrix(c("gndr", "iagpnt", "iagpnt_cat", "iagmr", "iagmr_cat",
"hinctnta", "hinctnta_cat", "hinctnta_cat_sorted", "eisced",
"eisced_cat", "eisced_cat_sorted", "admge", "admge_cat", "admge_cat2", 

"nominal", "ratio","ordinal", "ratio", "ordinal", "ordinal", "ratio",
"ordinal", "ordinal", "ordinal", "ordinal", "nominal", "nominal", "nominal",

"categorical", "continuous","categorical", "continuous", "categorical", "categorical", "continuous",
"categorical", "categorical","categorical","categorical","categorical","categorical","categorical",

class(ktn$gndr), class(ktn$iagpnt), class(ktn$iagpnt_cat), class(ktn$iagmr), class(ktn$iagmr_cat), class(ktn$hinctnta),
class(ktn$hinctnta_cat), class(cls1), class(ktn$eisced), class(ktn$eisced_cat), class(cls2),
class(ktn$admge), class(ktn$admge_cat), class(ktn$admge_cat2)), ncol = 4)

colnames(Table) <- c("Variables", "Measurement scale", "Variables’ types", "R data types")
Table <- as.data.frame(Table)
kbl(Table, align = "cccc") %>%
kable_styling(bootstrap_options = c("striped", "hover"))
```

## Analysis

Analysis of the first three hypotheses regarding gender differentiation is conducted through consideration of boxplots and histograms reflecting gender/subcategories and the ideal age for marriage or parenthood in the former with all mean and median values, and with mode values for the ideal age per category in the latter.  

### Hypothesis 1
#### Ideal age for getting married
```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x = admge_cat2, y = iagmr)) +
geom_boxplot()+
stat_summary(fun = "mean",colour = "#626FD9", size = 1)+
labs( x = "Gender", y = "Ideal age for getting married")
```

```{r warning = FALSE, message=FALSE}
ktn %>% 
  group_by(admge_cat2) %>% 
  summarise(mean = mean(iagmr, na.rm = T),
            median = median(iagmr, na.rm = T),
            mode = Mode(iagmr))
```

The horizontal lines along the boxplot represent the median values of the ideal age at marriage for each gender, while the circles reflect the mean values. We observe that both mean and median values are lower for women (around 25-26 years old) and higher for men (27-28 years old). 


```{r warning = FALSE, message=FALSE}
hist(f$iagmr,
main = "Women",
xlab="Ideal age for getting married")
abline(v = Mode(f$iagmr), col = '#626FD9', lwd = 3, lty = 2)

hist(m$iagmr,
main = "Men",
xlab="Ideal age for getting married")
abline(v = Mode(m$iagmr), col = '#626FD9', lwd = 3, lty = 2)
```

The vertical lines displayed in the histograms represent the mode values observed for each gender. The graphs indicate the presence of a similar trend as with the mean and median, since the mode value for men amounts to 30 and the histogram appears left-skewed, while the women's case shows a slightly reverse distribution with a mode value of 25.

Thus, a comparison of central tendency measures and distributions between men and women supports the first hypothesis, and both genders tend to report a lower age of marriage for women, which is consistent with the Todosijevic, Bojan, and Suzana Ignjatović (2021) findings.



#### Ideal age for becoming a parent 
``` {r warning = FALSE, message=FALSE}
 ggplot(ktn, aes(x = admge_cat2, y = iagpnt)) +
geom_boxplot()+
stat_summary(fun = "mean",colour = "#6F7CDF", size = 1)+
labs(x = "Gender", y = "Ideal age for becoming a parent")
```
```{r warning = FALSE, message=FALSE}
ktn %>% 
  group_by(admge_cat2) %>% 
  summarise(mean = mean(iagpnt, na.rm = T),
            median = median(iagpnt, na.rm = T),
            mode = Mode(iagpnt))
```

The horizontal lines along the boxplot represent the median values of the ideal age at marriage for each gender, while the circles reflect the mean values. The graph shows that respondents reported a lower appropriate age of parenthood for women, approximately 26 years for median and mean values ( though this age is higher than indicated within marriage for women), and higher for men, about 27-28.

```{r warning = FALSE, message=FALSE}
hist(f$iagpnt,
main = "Women",
xlab="Ideal age for becoming a parent")
abline(v = Mode(f$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)

hist(m$iagpnt,
main = "Men",
xlab="Ideal age for becoming a parent")
abline(v = Mode(m$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)
```

Mode values represent exactly the same situation as with marriage, with both a lower value and a slightly more right-skewed distribution for women, and a higher value with a left-skewing distortion for men. However, it should be mentioned that in general both distributions look a bit more symmetrical than with the ideal marriage age 
Accordingly, the first hypothesis is also confirmed for the ideal age of becoming a parent.

### Hypothesis 2 and 3
#### Ideal age for getting married
```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x = admge_cat, y = iagmr)) +
geom_boxplot()+
stat_summary(fun = "mean",colour = "#626FD9", size = 1)+
labs(x = "Category", y = "Ideal age for getting married")
```

The horizontal lines along the boxplot represent the median values of the ideal age at marriage for each of the four subcategories, whereas the circles reflect the mean values. We can detect that men and women reported approximately identical relevant and acceptable age of marriage for women (25-26), meanwhile, regarding the acceptable male age of marriage, men reported a lower age for themselves (about 26) than women did for them (about 28)

```{r warning = FALSE, message=FALSE}
hist(m1$iagmr,
main = "Men about women",
xlab="Ideal age for getting married")
abline(v = Mode(m1$iagmr), col = '#626FD9', lwd = 3, lty = 2)
hist(f1$iagmr,
main = "Women about women",
xlab="Ideal age for getting married")
abline(v = Mode(f1$iagmr), col = '#626FD9', lwd = 3, lty = 2)
```
```{r warning = FALSE, message=FALSE}
hist(m2$iagmr,
main = "Men about men",
xlab="Ideal age for getting married")
abline(v = Mode(m2$iagmr), col = '#626FD9', lwd = 3, lty = 2) 
hist(f2$iagmr,
main = "Women about men",
xlab="Ideal age for getting married")
abline(v = Mode(f2$iagmr), col = '#626FD9', lwd = 3, lty = 2)
```


Within the context of mode values, exactly the same pattern can be discerned, and the values most frequently marked by men and women for women coincide(25), while females often marked a higher age for men (30) than males for themselves (25). It can also be pointed out that women's answers concerning themselves take a more right-skewed shape, and those about men are more left-skewed, yet both graphs of men's answers appear to be more or less symmetrical.


```{r warning = FALSE, message=FALSE}
ktn %>% 
  group_by(admge_cat) %>% 
  summarise(mean = mean(iagmr, na.rm = T),
            median = median(iagmr, na.rm = T),
            mode = Mode(iagmr))
```

Therefore, based on the central tendency measures comparison, we can reject the second hypothesis, that men and women mark differently the ideal age of marriage for women, although still confirming the third hypothesis, stating that women report a higher age for men than they do for themselves.

#### Ideal age for becoming a parent 

```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x = admge_cat, y = iagpnt)) +
geom_boxplot()+
stat_summary(fun = "mean",colour = "#6F7CDF", size = 1)+
labs(x = "Category", y = "Ideal age for becoming a parent")
```

The horizontal lines along the boxplot represent the median values of the ideal age at marriage for each of the four subcategories, whereas the circles reflect the mean values. It can be noted that within the responses regarding the relative best age for women to become a parent, women tend to indicate a slightly higher age for themselves than men indicate for them within both the median and mean values. As for the ideal age for men, men clearly mark a lower acceptable age for themselves than women do indicate for them. 

```{r warning = FALSE, message=FALSE}
hist(m2$iagpnt,
main = "Men about men",
xlab="Ideal age for getting married")
abline(v = Mode(m2$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)
hist(f2$iagpnt,
main = "Women about men",
xlab="Ideal age for becoming a parent")
abline(v = Mode(f2$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)
```
```{r warning = FALSE, message=FALSE}
hist(m1$iagpnt,
main = "Men about women",
xlab="Ideal age for becoming a parent")
abline(v = Mode(m1$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)
hist(f1$iagpnt,
main = "Women about women",
xlab="Ideal age for becoming a parent")
abline(v = Mode(f1$iagpnt), col = '#6F7CDF', lwd = 3, lty = 2)
```

Mode values showed no differences in the opinions of men and women concerning one's and the opposite sex, and, accordingly, both sexes most often indicated the ideal age for having a child as 25 for a woman and 30 for a man. All of the histograms appear somewhat symmetrical.

```{r warning = FALSE, message=FALSE}
ktn %>% 
  group_by(admge_cat) %>% 
  summarise(mean = mean(iagpnt, na.rm = T),
            median = median(iagpnt, na.rm = T),
            mode = Mode(iagpnt))
```

Although the mode values did not show much difference, a comparison of the median and mean values supports the hypotheses according to which women report a higher age for men than they do for themselves, and that men report a lower age for women than they do for themselves.  This also corresponds with the findings of Todosijevic, Bojan, and Suzana Ignjatović (2021) in the framework of effects on reproduction related behavior

### Hypothesis 4
#### Educational level and ideal age for family formation 

```{r warning = FALSE, message=FALSE}
plot_xtab(ktn$eisced_cat_sorted , ktn$iagmr_cat, margin = "row", bar.pos = "stack")+
labs(x = "Level of education", y = "Frequency")+
scale_fill_manual(values = c("#434659","#A7AEDF" , "#626FD9" , "#4B55A6"),
name = "Ideal age for getting married",
labels = c("18 and earlier", "from 19 to 25", "from 26 to 30", "over 31"))
```

The stacked bar plot indicates the frequency of choosing the ideal age for marriage in the 4 timing categories, depending on the level of education. Hence,  we can see that people with a low level of education tend to indicate a young ideal age for getting married (19-25 years) more frequently. The share of the people who think so is declining. People with higher education less often choose this period of life as ideal for creating a family. The proportion of people indicating 25-30 years of age increases with the level of education

```{r warning = FALSE, message=FALSE}
ggplot(na.omit(ktn), aes(x = eisced_cat_sorted, y = iagmr)) +
  geom_boxplot() +
  stat_summary(fun = "mean",colour = "#626FD9", size = 1)+
  labs(x = "Level of education", y = "Ideal age for getting married")

```

From the mean values (pink points on the graph) we can trace a similar trend, with the high education group having the highest value of perceiving the ideal age of marriage.

```{r warning = FALSE, message=FALSE}
plot_xtab(ktn$eisced_cat_sorted, ktn$iagpnt_cat, margin = "row", bar.pos = "stack") +
  labs(x = "Level of education", y = "Frequency") +scale_fill_manual(values = c("#434659","#A7AEDF" , "#626FD9" , "#4B55A6"),
name = "Ideal age for becoming a parent",
labels = c("18 and earlier", "from 19 to 25", "from 26 to 30", "over 31"))

```

The stacked bar plot indicates the frequency of choosing the ideal age for becoming a parent in the 4 timing categories, depending on the level of education. We can confirm the hypothesis that the ideal age for becoming parents increases as the level of education increases: the proportion of those with higher education most frequently chose the ideal age for becoming parents from 26 to 30 and from 31+, while the main periods for middle and low educational levels were from 19 to 25 years and below.

```{r warning = FALSE, message=FALSE}
ggplot(na.omit(ktn), aes(x = eisced_cat_sorted, y = iagpnt)) +
  geom_boxplot() +
  stat_summary(fun = "mean",colour = "#626FD9", size = 1)+
  labs(x = "Level of education", y = "Ideal age for becoming a parent")
```

Again, from the mean values (pink points on the graph) we can trace a similar trend, with the high education group having the highest value of perceiving the ideal age of becoming a parent.


### Hypothesis 5 
#### Income and ideal age for family formation 

```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x = hinctnta_cat, y = iagmr )) +
  geom_point() +
  geom_smooth(method="lm")+
  labs(x = "Income decile", y = "Ideal age for getting married")
```

The scatter plot represents individuals' responses on the ideal time of entering a marriage, depending on their income background (in deciles).
According to the results, we see that the level of income has negligible influence on determining the ideal age for marriage, nethertheless we observe an upward trend 

```{r warning = FALSE, message=FALSE}
ggplot(ktn, aes(x = hinctnta_cat, y = iagpnt )) +
  geom_point() +
  geom_smooth(method="lm")+ 
 labs(x = "Income decile", y = "Ideal age for becoming a parent")
```

The scatter plot represents individuals' responses on the ideal time of becoming a parent, depending on their income background (in deciles).
According to the results, we see that the level of income has a similar minor influence on determining the ideal age for becoming a parent as well, however the slope is generally positive

## Discussion and conclusions

The purpose of this study was to examine how demographic (gender) and socioeconomic (income and education) factors impact the perceived ideal timing of family formation in individuals. 

Under the influence of gender on the perceived age of marriage, we were able to confirm the first and third hypotheses for the mean, median and mode values, suggesting that women are generally ascribed a lower ideal marital age, as substantiated by evolutionary theory, and that women estimate the ideal age of marriage for men to be higher than men for themselves, as they are concerned about their spouse being able to support the family. The second hypothesis was rejected, as women and men estimated the ideal age of marriage entry for females similarly.

Considering the effect of gender on the perceived ideal age of becoming a parent, all three hypotheses were confirmed within the median and average values.
Regarding the effect of income and education on the perceived ideal timing of family formation, we can say that people with higher education prefer a later timing, thus confirming our 4th hypothesis, but we did not find a strong relationship between income and timing for our 5th hypothesis, but a positive trend does occur.


# Chi-square, T-test, ANOVA analysis

#### Hypothesis 1

We hypothesize that there would be statistically significant association between respondent's education level and perceived ideal age for getting married, that is, there will be a greater proportion of respondents with higher level of education (compared to lower one) who will mark a higher ideal age for getting married.

#### Hypothesis 2

We hypothesize that there would be a statistically significant association between perceived ideal age to get married and gender answer group , such that a greater proportion of men (compared to women) would indicate a lower ideal age to get married for women, and conversely a greater proportion of women (compared to men) will indicate a higher ideal age to get married for men.

#### Hypothesis 3

We hypothesize that means of perceived ideal age to get married will be lower for women than for men from evaluations by both genders.

#### Hypothesis 4

We hypothesize that there will be a difference in means of perceived ideal age for getting married, according to respondents' income level, that is, respondents with higher income, on average, will mark higher age for getting married, than respondents with lower level of income.

## Variables

Prior to forming all the necessary categories and manipulating variables, four lines were removed from the data, the values of which on the main issues for us (the ideal age for marriage ) were either missing or were irrelevant within the family formation (such values as 2 for marriage).

```{r warning = FALSE, message=FALSE}
ktn$iagmr <- as.numeric(as.character(ktn$iagmr))

# Ideal age for marriage
ktn$iagmr_cat[ktn$iagmr <="18"] <- "18 and earlier"
ktn$iagmr_cat[ktn$iagmr >= "19"& ktn$iagmr <="25"] <- "from 19 to 25"

ktn$iagmr_cat[ktn$iagmr >= "26"& ktn$iagmr <"30"] <- "from 26 to 30"
ktn$iagmr_cat[ktn$iagmr >= "31"] <- "over 31"

```

Consequently, comparison of the reported ideal age of getting married occurred according to the 4 main criterion of respondents:

### Gender

The gender subgroups are evenly distributed into two categories, amounting to 778 for men and 757 for women, respectively.


### Subgroup on the question (response about females or males)

```{r warning = FALSE, message=FALSE}
#Gender subgroups datasets
f1 <- dplyr::filter(ktn, ktn$admge == "Group 1" & ktn$gndr == "Female")
m1 <- dplyr::filter(ktn, ktn$admge == "Group 1" & ktn$gndr == "Male")
f2 <- dplyr::filter(ktn, ktn$admge == "Group 2" & ktn$gndr == "Female")
m2 <- dplyr::filter(ktn, ktn$admge == "Group 2" & ktn$gndr == "Male")

f <- dplyr::filter(ktn, ktn$admge == "Group 1")
m <- dplyr::filter(ktn, ktn$admge == "Group 2")

#Gender subgroups 
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Female"] <- "Women about Women"
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Male"] <- "Men about Women"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Female"] <- "Women about Men"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Male"] <- "Men about Men"

ktn$admge_cat2[ktn$admge == "Group 1"] <- "About Women"
ktn$admge_cat2[ktn$admge == "Group 2"] <- "About Men"

```

By considering which gender answer subgroups the person is in (group 1 or group 2) and comparing this to the gender of the person him/herself, we generated 4 subgroups, likewise evenly distributed, comprising about 380-400 people, with the category of women about men being the lowest with 357 people

### Income

```{r warning = FALSE, message=FALSE}
# Income decile
ktn$hinctnta_cat[ktn$hinctnta == "J - 1st decile"] <- "1"
ktn$hinctnta_cat[ktn$hinctnta == "R - 2nd decile"] <- "2"
ktn$hinctnta_cat[ktn$hinctnta == "C - 3rd decile"] <- "3"
ktn$hinctnta_cat[ktn$hinctnta == "M - 4th decile"] <- "4"
ktn$hinctnta_cat[ktn$hinctnta == "F - 5th decile"] <- "5"
ktn$hinctnta_cat[ktn$hinctnta == "S - 6th decile"] <- "6"
ktn$hinctnta_cat[ktn$hinctnta == "K - 7th decile"] <- "7"
ktn$hinctnta_cat[ktn$hinctnta == "P - 8th decile"] <- "8"
ktn$hinctnta_cat[ktn$hinctnta == "D - 9th decile"] <- "9"
ktn$hinctnta_cat[ktn$hinctnta == "H - 10th decile"] <- "10"



ktn$hinctnta_cat <- as.numeric(ktn$hinctnta_cat)

ktn$hinctnta_cat_1[ktn$hinctnta_cat == "1" | ktn$hinctnta_cat == "2" ] <- "low income" 
ktn$hinctnta_cat_1[ktn$hinctnta_cat == "3" | ktn$hinctnta_cat == "4" | ktn$hinctnta_cat == "5" ] <- "below average" 
ktn$hinctnta_cat_1[ktn$hinctnta_cat == "6" | ktn$hinctnta_cat == "7" | ktn$hinctnta_cat == "8"] <- "above average" 
ktn$hinctnta_cat_1[ktn$hinctnta_cat == "9" | ktn$hinctnta_cat == "10"] <- "high income"

ktn$hinctnta_cat_1_sorted <- factor(ktn$hinctnta_cat_1, ordered = TRUE, levels = c("low income" , "below average", "above average", "high income"))


ktn$hinctnta_cat_1 <- factor(ktn$hinctnta_cat_1)


```

Respondents' incomes are compiled using the decile income distribution from the original database.  Individuals are arranged in ascending order of their income and divided into ten groups of equal size. Each income decile contains 10 percent of the population; the higher the decile, the higher the income level.



### Education

```{r warning = FALSE, message=FALSE}
# education levels
ktn$eisced_cat[ktn$eisced == "ES-ISCED I , less than lower secondary"|
                 ktn$eisced == "ES-ISCED II, lower secondary"] <- "low education"
ktn$eisced_cat[ktn$eisced =="ES-ISCED IIIa, upper tier upper secondary" |
                 ktn$eisced =="ES-ISCED IIIb, lower tier upper secondary"|
                 ktn$eisced =="ES-ISCED IV, advanced vocational, sub-degree"] <- "middle education"
ktn$eisced_cat[ktn$eisced == "ES-ISCED V1, lower tertiary education, BA level" |
                 ktn$eisced == "ES-ISCED V2, higher tertiary education, >= MA level"] <- "high education"

ktn$eisced_cat_sorted <- factor(ktn$eisced_cat, ordered = TRUE,
levels = c("low education", "middle education", "high education"))
```

The level of education has been distributed according to International Standard Classification of Education, where ISCED levels 0-2 correspond to low education, ISCED levels 3-4 correspond to middle education, ISCED levels 5-8 correspond to higher education

Our data ended up having 1535 observations and 14 variables, 6 of which were taken directly from the ESS9 database, while the other 8 were created.
``` {r warning = FALSE, message=FALSE}
cls1 <- class(ktn$hinctnta_cat_sorted)
cls2 <- class(ktn$eisced_cat_sorted)
Table <- matrix(c("gndr", "iagmr", "iagmr_cat",
"hinctnta", "hinctnta_cat", "hinctnta_cat_sorted", "hinctnta_cat_1", "eisced",
"eisced_cat", "eisced_cat_sorted", "admge", "admge_cat", "admge_cat2", 

"nominal", "ratio", "ordinal", "ordinal", "ratio",
"ordinal", "ordinal",  "ordinal", "ordinal", "ordinal", "nominal", "nominal", "nominal",

"categorical",  "continuous", "categorical", "categorical", "continuous",
"categorical", 'catgorical', "categorical","categorical","categorical","categorical","categorical","categorical",

class(ktn$gndr), class(ktn$iagmr), class(ktn$iagmr_cat), class(ktn$hinctnta),
class(ktn$hinctnta_cat), class(cls1), class(ktn$hinctnta_cat_1), class(ktn$eisced), class(ktn$eisced_cat), class(cls2),
class(ktn$admge), class(ktn$admge_cat), class(ktn$admge_cat2)), ncol = 4)

colnames(Table) <- c("Variables", "Measurement scale", "Variables’ types", "R data types")
Table <- as.data.frame(Table)
kbl(Table, align = "cccc") %>%
kable_styling(bootstrap_options = c("striped", "hover"))
```

## Chi-square Test

### Education & Ideal age of marriage 

We want to explore the relationship between two categorical variables of education level (eisced_cat_sorted) and ideal age of marriage (iagmr_cat).

```{r warning = FALSE, message=FALSE}
plot_xtab(ktn$eisced_cat_sorted , ktn$iagmr_cat, margin = "row", bar.pos = "stack")+
labs(x = "Level of education", y = "Frequency")+
scale_fill_manual(values = c("#434659","#A7AEDF" , "#626FD9" , "#4B55A6"),
name = "Ideal age for getting married",
labels = c("18 and earlier", "from 19 to 25", "from 26 to 30", "over 31"))

```

#### Contingency table

```{r warning = FALSE, message=FALSE}
table(ktn$iagmr_cat, ktn$eisced_cat_sorted)
```

#### Null and Alternative Hypotheses

**Null hypothesis (H0)**:  the variables of ideal age of marriage and level of education are independent 

**Alternative hypothesis (H1)**: the variables of ideal age of marriage and level of education are dependent


```{r warning = FALSE, message=FALSE}

# Chi-square Test
chisq.test(ktn$iagmr_cat, ktn$eisced_cat_sorted)
s <- chisq.test(ktn$iagmr_cat, ktn$eisced_cat_sorted)
```
There is a statistically significant association between level of education and  ideal age of marriage indicated by respondent, X-squared(6) = 24.398, p-value = 0.0004412, so the variables are not independently distributed. It means that the level of education of the respondent is related to the perceived ideal age to get married.

Now we are going to check the chi-square test assumption on having at least 5 expected observations per cell and then inspect and visualize the standardized residuals to understand which categories create the largest differences between the model and the data.

```{r warning = FALSE, message=FALSE}
s$expected
```
There are no expected frequencies below 5, so the assumption is met (the test is reliable).

#### Analyzing Standardized residuals

```{r warning = FALSE, message=FALSE}
#STANDARDIZED RESIDUALS

col<- colorRampPalette(c("deeppink3", 'gainsboro',"darkseagreen3"))(20)
corrplot(s$residuals, is.cor = FALSE, col = col)
s$stdres
```
We can conclude that respondents with low education are more likely to indicate period of 19 to 25 years for ideal age to get married, and less likely to state the period over 31 than the model would expect.
As well, there are less respondents with high education that have chosen the period of 19 to 25 years for ideal marriage age, and much more people who stated the period from 26 to 30 than we would expect.

Finally, we would like to plot the association between variables and inspect how strong it is using the Cramér’s V measure.  
```{r warning = FALSE, message=FALSE}
#EFFECT SIZE
assocstats(table(ktn$eisced_cat_sorted , ktn$iagmr_cat))

# PLOTS
ggbarstats(
  data = ktn,
  x = iagmr_cat,
  y = eisced_cat_sorted,
  legend.title = 'Ideal age for getting married'
) +
  labs(x = 'Level of education') + 
 scale_fill_manual(values = c("#E2ADF2","#A7AEDF" , "#AB81CD", "#222A68" ),
labels = c("18 and earlier", "from 19 to 25", "from 26 to 30", "over 31"))
```

Cramer’s V of 0.115 at degrees of freedom = 6 indicates that although the result is statistically significant, the fields are only weakly associated.

Concluding, we can confirm our first research hypothesis, and state that there is an association, though it is small, between variables of education level and ideal age for marriage and residuals verify that more respondents from high education group are indicating higher values of perceived ideal age of getting married, whereas more of those from lower education one have chosen lower age as ideal.

### Gender answer subgroup & Ideal age of marriage 

We want to explore the relationship between two categorical variables of gender answer subgroup (admge_cat) and ideal age of marriage (iagmr_cat).

```{r warning = FALSE, message=FALSE}
plot_xtab(ktn$admge_cat , ktn$iagmr_cat, margin = "row", bar.pos = "stack")+
labs(x = "Level of education", y = "Frequency")+
scale_fill_manual(values = c("#434659","#A7AEDF" , "#626FD9" , "#4B55A6"),
name = "Ideal age for getting married",
labels = c("18 and earlier", "from 19 to 25", "from 26 to 30", "over 31"))

```

#### Contingency table

```{r warning = FALSE, message=FALSE}
table(ktn$iagmr_cat, ktn$admge_cat)
```

#### Null and Alternative Hypotheses


**Null hypothesis (H0)**: the variables of ideal age of marriage and gender answer subgroup are independent 


**Alternative hypothesis (H1)**: the variables of ideal age of marriage and gender answer subgroup are dependent

```{r}
chisq.test(ktn$iagmr_cat, ktn$admge_cat)
k <- chisq.test(ktn$iagmr_cat, ktn$admge_cat)
```
There is a statistically significant association between gender answer subgroup and  ideal age of marriage indicated by respondent, X-squared(9) = 27.3, p-value = 0.001248, so the variables are not independently distributed. It means that gender answer subgroup(about men or about women) of the respondent is related to the perceived ideal age to get married .

Now we are going to check the chi-square test assumption on having at least 5 expected observations per cell and then inspect and visualize the standardized residuals to understand which categories create the largest differences between the model and the data.

```{r warning = FALSE, message=FALSE}
k$expected
```
There are no expected frequencies below 5, so the assumption is met (the test is reliable).

#### Analyzing Standardized residuals

```{r warning = FALSE, message=FALSE}

col<- colorRampPalette(c("deeppink3", 'gainsboro',"darkseagreen3"))(20)
corrplot(k$residuals, is.cor = FALSE, col = col)
k$stdres
```
We can conclude that male respondents giving answer about women are more likely to indicate the period of 18 and earlier as ideal age to get married for them than the model would expect. 
As well, there are much less female respondents giving answer about men that have chosen the period of 18 and earlier as ideal marriage age for them, and much more of those who have chosen the period of over 31, than we would expect. In addition, female respondents are less likely to state the period of over 31 years as ideal marriage age for themselves. 

Finally, we would like to plot the association between variables and inspect how strong it is using the Cramér’s V measure.  

```{r warning = FALSE, message=FALSE}

#EFFECT SIZE
assocstats(table(ktn$admge_cat , ktn$iagmr_cat))

# PLOTS
ggbarstats(
  data = ktn,
  x = iagmr_cat,
  y = admge_cat,
  legend.title = 'Ideal age for getting married'
) +
  labs(x = 'Gender answer subgroup') + 
  scale_fill_manual(values = c("#E2ADF2","#A7AEDF" , "#AB81CD", "#222A68" ))

```

Cramer’s V of 0.098 at degrees of freedom = 9 indicates that although the result is statistically significant, the fields are very weakly associated.

Thus, we can as well confirm our second research hypothesis, and conclude that there is an association, though it is weak, between variables of gender answer subgroup and ideal age for marriage and residuals verify that more male respondents are indicating lower values of perceived ideal age of getting married for women, whereas more female respondents have chosen higher ideal age for men.

## T-Test 

#### Answer group  & Ideal age of marriage 

We want to explore the relationship between the categorical variable of gender answer subgroup (admge_cat2) and a continuous variable of ideal marriage age(iagmr).

```{r warning = FALSE, message=FALSE}

ggplot(ktn, aes(x = admge_cat2)) + geom_bar(fill = "#A7AEDF") + coord_flip() + scale_x_discrete(limits = rev(levels(ktn$admge_cat)))+
labs(x = "Categories", y = "Number of people")

ggplot(ktn, aes(iagmr)) + 
  geom_histogram(color = "#ffffff" , fill = "#A7AEDF", 
                 bins = 15)+
  labs(x = 'Ideal age for getting married')
```

**Null hypothesis (H0)**:  there is no difference between mean ideal age of marriage for gender answer subgroups (about men or about women) 

**Alternative hypothesis (H1)**: mean ideal age of marriage for gender answer subgroups (about men or about women)  are different 



### Checking Assumptions
#### Normality of the variable distribution
```{r warning = FALSE, message=FALSE}
#NORAMILTY
# HISTOGRAM
mu <- ddply(ktn, "admge_cat2", summarise, grp.mean = mean(iagmr, na.rm = T)) # calculate mean by group 

ggplot(ktn,aes(x = iagmr, color = admge_cat2, fill = admge_cat2)) +geom_density(alpha = 0.5) +
  labs(title = "Answers by gender", x = "Ideal age for marriage", y = "Density") +theme_classic()

# QQ-plot
par(mfrow =c(1,2))
qqnorm(f$iagmr);qqline(f$iagmr, col = 2)
qqnorm(m$iagmr);qqline(m$iagmr, col = 2)
```

According to histogram and qq-plot, we observe a non-normal distribution. However, the number of observations is large (n>100), and the t-test is highly robust to non-normality, so we can proceed with the t-test and additionally compare received results with a non-parametric test.

#### Homogenity of variances
To check the homogenity of variances assumption we will proceed with the Levene's test as it is less sensittive to departures from normality.

```{r warning = FALSE, message=FALSE}
# HOMOGENITY OF VARIANCES
leveneTest(ktn$iagmr ~ ktn$admge_cat2)
```
Levene’s test showed big p-value (>0,05), so we cannot reject the null hypothesis. Thus, the variances are equal. 


### Applying T-test

The distributions of the DV are not normal but the sample is large enough (n > 100), so non-normality can be effectively ignored. Variances are equal, so the classic T-test can be applied. 

```{r warning = FALSE, message=FALSE}
t.test(ktn$iagmr ~ ktn$admge_cat2)
```
We can conclude that:

1. On average, the ideal age of marriage for men is 27.4 and for women 26.1
2. The t-statistic t(1312) =  6.2049 (p-value < 0.001), which means that the observed difference in means is statistically significant across the two subgroups, answering about different genders (higher in group “about men”).

### Non-parametic T-test

Now, as our distributions are not normal, we would also like to do a non-parametric test in order to compare the results.
```{r}
wilcox.test(iagmr ~ admge_cat2, data = ktn)
```

The Wilcoxon rank sum W equals 255593 at p < 0.001, which means that two subgroups, answering about different genders, perceive the ideal age for getting married differently and this difference is statistically significant. 

### Effect size

Finally, we would like to explore the effect size - how large the observed difference actually is.
```{r}
cohen.d(ktn$iagmr ~ ktn$admge_cat2, na.rm = T)
rstatix::wilcox_effsize(iagmr ~ admge_cat2, data = ktn, na.rm = T) # for non-parametric
```

With a Cohen's d of 0.34, 63.3% of the group answering about men will be above the mean of the group answering about women (Cohen's U3), 86.5% of the two groups will overlap, and there is a 59.5% chance that a person picked at random from the group answering about men will have a higher score than a person picked at random from the group answering about women (probability of superiority). Moreover, in order to have one more favorable outcome in the group answering about men compared to the group answering about women, we need to treat 9.3 people on average. This means that if there are 100 people in each group, and we assume that 20 people have favorable outcomes in the group answering about women, then 20 + 10.8 people in the group answering about men will have favorable outcomes.

The effect size of the Mann-Whitney-Wilcoxon test is 0,16 (small), even though it is statistically significant. 

Concluding, we can confirm our third research hypothesis and there is a difference between the ideal age for getting married between the answering groups, even though it is small, and both genders indicate a higher ideal marriage age for men and a lower one for women.

## ANOVA 

We want to explore the relationship between the categorical variable of income levels (hinctnta_cat_1) and a continuous variable of ideal age of marriage (iagmr), to assess whether the mean of the ideal age outcome variable is different across 4 groups.


```{r warning = FALSE, message=FALSE}
# Ideal age to get married
ggplot(ktn, aes(x = iagmr)) + 
geom_density( 
color="#4B55A6", 
fill="#A7AEDF",)+
  labs(x = 'Ideal age for getting married')

# Income category
par(mar = c(5, 10, 2, 2))
barplot(table(ktn$hinctnta_cat_1_sorted)/nrow(ktn)*100, las = 2, horiz = T, col = "#A7AEDF", xlim = c(0, 60))

# Boxplot 
ggboxplot(ktn, x = 'hinctnta_cat_1', y = 'iagmr', 
          color = 'hinctnta_cat_1', palette = c("#E2ADF2","#A7AEDF" , "#AB81CD", "#222A68" ),
          order = c("low income", "below average", "above average", "high income"),
          ylab = "Ideal age to get married", xlab = "Income group", legend.title = 'Income group')
    


```


**Null hypothesis (H0)**: The Mean Ideal ages for getting married indicated by all groups are the same

**Alternative hypothesis (H1)**: At least one sample Mean Ideal age for getting married is different

### Values descriptives

```{r warning = FALSE, message=FALSE}

describeBy(ktn$iagmr, ktn$hinctnta_cat_1_sorted, mat = TRUE) %>%#create dataframe
select(Income = group1, N=n, Mean=mean, SD=sd, Median=median, Min=min, Max=max,                Skew=skew, Kurtosis=kurtosis, st.error = se) %>%
kable(align=c("lrrrrrrrr"), digits=2, row.names = FALSE,        caption="Ideal age for marriage by Income") %>%
kable_styling(bootstrap_options=c("bordered", "responsive","striped"), full_width = FALSE)
```

### Applying ANOVA test

#### Homogenity of variances

```{r warning = FALSE, message=FALSE}
leveneTest(ktn$iagmr ~ ktn$hinctnta_cat_1)

```

From the output we can see that the p-value is equal to 0,56, so there is no evidence to suggest that the variance across groups are different. Therefore, we can assume the homogeneity of variances in different income groups.

#### one-way ANOVA test

```{r warning = FALSE, message=FALSE}
aov.out <- aov(ktn$iagmr ~ ktn$hinctnta_cat_1)
summary(aov.out)

```

F(3, 1237) = 9,436, p-value < 0.001 means that there is statistically significant difference in the indicated ideal age for getting married across income groups.


#### Normality of residuals
Now we would like to check the assumption that the residuals are normally distributed. It should approximately follow a straight line.
```{r warning = FALSE, message=FALSE}

# QQ plot
plot(aov.out, 2, sub = '')
```

The data points lie mostly along the diagonal line, which means that the distribution of residuals is close to normal.

```{r}
# skew and kurtosis
anova.res <-residuals(object = aov.out)
describe(anova.res)
# Histogram
hist(residuals(aov.out))

#Formal test
shapiro.test(residuals(aov.out))
```


Skew(0.17) < 0.5 and kurtosis(1.65) < 2 depict normal distribution.

However, the Shapiro-Wilk test says that it is not normal, but at larger sample sizes the Shapiro-Wilk test becomes very sensitive even to a minor deviation from normality.
Visual analysis and skew and kurtosis agree here, which is why we can conclude that the normality assumption holds. 

### Post hoc test
A significant p-value has shown that some of the group means are different, but we don’t know yet which pairs of groups are different, so we would like to perform multiple pairwise-comparison, to determine which groups contribute to the statistical significance of this test.

```{r warning = FALSE, message=FALSE}
# post hoc test

TukeyHSD(aov.out)

par(mar =c(5, 15, 3, 1))
Tukey <-TukeyHSD(aov.out)
plot(Tukey, las = 2)

ggbetweenstats(ktn, x = hinctnta_cat_1_sorted, y = iagmr,
               plot.type = "box", # we believe it looks better than the default version
               type = "parametric",
               ylab = 'Ideal age for getting married',
               xlab = 'Income groups')

```

We can see that differences between three pairs of groups are statistically significant.

Difference between below average income group and above average income group is significant with an adjusted p-value of 0.0079636.

Difference between below average income group and high income group is significant with an adjusted p-value of 0.0000187.

Difference between low income group and high average income group is significant with an adjusted p-value of 0.0007861.

### Effect size
Finally, we want to establish what is the substantive significance of income level in determining the ideal age for getting married.

```{r warning = FALSE, message=FALSE}
anova_stats(aov.out) 
```
The omega-squared is 0.020 - this is a low effect. It means that even though there is a statistically significant difference in the perceived ideal age of getting married across income groups and this difference is statistically significant across all but three pairs of groups, in practical terms this effect is modest.

### Non-parametric ANOVA test
Now we would also like to do a non-parametric test in order to double-check and compare the results.

```{r warning = FALSE, message=FALSE}
kruskal.test(iagmr ~ hinctnta_cat_1, data = ktn)

```
With KW chi-square(3) = 34.321, p-value is < .001, which means that the mean ranks of the income groups are different. This results confirms what we saw earlier in the ANOVA test.

In order to determine which pairs of group contributed to the significance of the test to most we will run Dunn’s test.
```{r warning = FALSE, message=FALSE}
DunnTest(iagmr ~ hinctnta_cat_1, data = ktn)

ggbetweenstats(ktn, x = hinctnta_cat_1_sorted, y = iagmr,
               plot.type = "box", # we believe it looks better than the default version
               type = "non- parametric",
               ylab = 'Ideal age for getting married',
               xlab = 'Income groups')
```
These results show that all but one pair of groups have statistically significant differences in their medians.

### Effect size
Now, we would like to explore the effect size of the observed differences.

```{r warning = FALSE, message=FALSE}
kruskal_effsize(data = ktn, iagmr ~ hinctnta_cat_1)

```
The effect size of the non-parametric test confirms that the effect is significant, although small.

Thus, we can confirm our 4th research hypothesis as there is statistically significant difference between indicated ideal ages for getting married for different income levels, and the higher the income the higher the chosen ideal marriage age.

## Discussion and conclusions

The purpose of this study was to examine how demographic (gender) and socioeconomic (income and education) factors impact the perceived ideal timing of getting married for individuals. 

Under the influence of gender on the perceived age of marriage, we were able to confirm the second and third research hypotheses through testing the presence of statistically significant association between gender answer subgroups and chosen ideal marriage age and whether there is any difference in mean values indicated for women and men by respondents.

Results suggested that women are generally ascribed a lower ideal marital age, as substantiated by evolutionary theory, and that women estimate higher ideal age of marriage for men, as they are concerned about their spouse being able to support the family, and men generally chose a lower ideal marriage age for women, being driven by ideas about their reproductive value and desire to date younger women. 

Regarding the effect of income and education on the perceived ideal timing of getting married, we were able to again confirm both of our hypotheses, by checking the presence of association between education levels and ideal age for marriage and through comparison of mean values across income group. We can say that bigger proportion of people with higher education preferred a later timing of marriage, while low education respondents, in contrary, were more prone to chose lower age, thus confirming our 1st hypothesis. And the income groups did show significant differences in the mean marriage age values as well, which have increased as the income got higher. 

# Correlation and regression analysis

#### Hypothesis 1

We hypothesize that there would be statistically significant association between respondent's income and perceived ideal age for getting married, such that greater income on average will lead to higher indicated ideal marriage age.

#### Hypothesis 2

We hypothesize that there would be statistically significant association between the level of education and perceived ideal age for getting married, such that higher level of education will result in higher ideal age for getting married as well.

#### Hypothesis 3

We hypothesize that there would be statistically significant association between the years of full-time education and perceived ideal age for getting married, such that the greater number of years the person has spent on education will result in higher ideal age for getting married as well.

#### Hypothesis 4

We hypothesize that there would be a statistically significant association between perceived ideal age to get married and gender answer group, such that men (compared to women) will indicate a lower ideal age to get married for women, and conversely women (compared to men) will indicate a higher ideal age to get married for men.

#### Hypothesis 5

We hypothesize that there would be statistically significant association between perceived ideal age for becoming a parent and perceived ideal age for getting married, such that indicated ages for these two events will be marked within roughly the same timeframe and higher ideal parenting age will be associated with higher ideal marriage age.

#### Hypothesis 6

We hypothesize that there would be statistically significant association between perceived ideal age for living with a partner and perceived ideal age for getting married, such that indicated ages for these two events will be marked within roughly the same timeframe as well and higher ideal cohabitation age will be associated with higher ideal marriage age, since cohabitation is being usually perceived as the pre-phase of marriage.

#### Hypothesis 7

We hypothesize that there would be statistically significant association between the level of religiosity and perceived ideal age for getting married, such that religious people will indicate a lower age for getting married.

#### Hypothesis 8

We hypothesize that there would be statistically significant association between respondents' age and perceived ideal age for getting married, such that the higher the age of the respondent, the lower the age indicated as ideal for marriage.

## Variables

Prior to forming all the necessary categories and manipulating variables, three lines were removed from the data, the values of which on the main issue for us (the ideal age for marriage) were either missing or were irrelevant within the family formation (such values as age "2" for marriage).

#### Ideal age to get married

```{r warning = FALSE, message=FALSE}
ktn$iagmr <- as.numeric(as.character(ktn$iagmr))
ggplot(ktn, aes(iagmr)) + 
  geom_histogram(fill = "#A7AEDF", 
                 bins = 15)+labs(x = 'Ideal age for getting married')
```

```{r warning = FALSE, message=FALSE}
summary(ktn$iagmr)
```

The distribution of respondents' answers regarding the ideal age of marriage is non-normal, with a median value of 27 and a mean of 26.78.

#### Ideal age to become a parent

```{r warning = FALSE, message=FALSE}
ktn$iagpnt <- as.numeric(as.character(ktn$iagpnt))
ggplot(ktn, aes(iagpnt)) + 
  geom_histogram(fill = "#A7AEDF", 
                 bins = 15)+labs(x = 'Ideal age to become a parent')
```

```{r warning = FALSE, message=FALSE}
summary(ktn$iagpnt)
```

The distribution of respondents' answers regarding the ideal age of becoming a parent is similarly non-normal, though the median and mean values are higher than for marriage one. They are 28 and 27.38, respectively.

#### Ideal age to start living with a partner

```{r warning = FALSE, message=FALSE}

ktn$iaglptn <- as.numeric(as.character(ktn$iaglptn))
#Plot
ggplot(ktn, aes(iaglptn)) + 
  geom_histogram(fill = "#A7AEDF", 
                 bins = 15)+labs(x = 'Ideal age to start living with a partner')
#Descriptive Statistics
summary(ktn$iaglptn)
```

Respondents' answers regarding the ideal age to start living with a partner are distributed pretty normally, and mean and median values equal 22 years.

#### Age

```{r warning = FALSE, message=FALSE}
ktn$yrbrn <- as.numeric(as.character(ktn$yrbrn))
ktn$age <- (2018-ktn$yrbrn)
ktn$age <- as.numeric(ktn$age)

#Plot
ggplot(ktn, aes(age)) + 
  geom_histogram(fill = "#A7AEDF", 
                 bins = 15)+labs(x = 'Age')
#Descriptive Statistics
summary(ktn$age)
```

To construct the age variable we have subtracted the respondents year of birth from the year of constructing the survey (2018). The distribution of respondents' age looks not really normal, with the median of 54 and mean age of 52.

#### Income

```{r warning = FALSE, message=FALSE}
# Income decile
ktn$hinctnta_cat[ktn$hinctnta == "J - 1st decile"] <- "1"
ktn$hinctnta_cat[ktn$hinctnta == "R - 2nd decile"] <- "2"
ktn$hinctnta_cat[ktn$hinctnta == "C - 3rd decile"] <- "3"
ktn$hinctnta_cat[ktn$hinctnta == "M - 4th decile"] <- "4"
ktn$hinctnta_cat[ktn$hinctnta == "F - 5th decile"] <- "5"
ktn$hinctnta_cat[ktn$hinctnta == "S - 6th decile"] <- "6"
ktn$hinctnta_cat[ktn$hinctnta == "K - 7th decile"] <- "7"
ktn$hinctnta_cat[ktn$hinctnta == "P - 8th decile"] <- "8"
ktn$hinctnta_cat[ktn$hinctnta == "D - 9th decile"] <- "9"
ktn$hinctnta_cat[ktn$hinctnta == "H - 10th decile"] <- "10"

ktn$hinctnta_cat <- as.numeric(ktn$hinctnta_cat)
#Plot
ggplot(na.omit(ktn), aes(x = hinctnta_cat)) + geom_bar(fill = "#A7AEDF") +
labs(x = "Income decile", y = "Number of people")
#Descriptive Statistics
summary(ktn$hinctnta_cat)
```

Respondents' incomes are compiled using the decile income distribution from the original database. Individuals are arranged in ascending order of their income and divided into ten groups. Each income decile contains 10 percent of the population; the higher the decile, the higher the income level. The mean and median income value among our respondents equals 6th decile.

#### Education

```{r warning = FALSE, message=FALSE}
ktn$eisced_cat[ktn$eisced == "ES-ISCED I , less than lower secondary"|
                 ktn$eisced == "ES-ISCED II, lower secondary"] <- "low education"
ktn$eisced_cat[ktn$eisced =="ES-ISCED IV, advanced vocational, sub-degree"|ktn$eisced =="ES-ISCED IIIa, upper tier upper secondary" |
                 ktn$eisced =="ES-ISCED IIIb, lower tier upper secondary"] <- "middle education"
ktn$eisced_cat[ktn$eisced == "ES-ISCED V1, lower tertiary education, BA level"|ktn$eisced == "ES-ISCED V2, higher tertiary education, >= MA level"] <- "high education"

ktn$eisced_cat_sorted <- factor(ktn$eisced_cat, ordered = TRUE,
levels = c("low education", "middle education", "high education"))
#Plot
ggplot(na.omit(ktn), aes(x =eisced_cat_sorted)) + geom_bar(fill = "#A7AEDF")+
  labs(x="Level of education", y = "Number of people")
```

The level of education has been distributed according to International Standard Classification of Education, where ISCED levels 0-2 correspond to low education, ISCED levels 3-4 correspond to middle education, ISCED levels 5-8 correspond to higher education.

```{r warning = FALSE, message=FALSE}
ktn$eduyrs <-as.numeric(ktn$eduyrs)
#Plot
ggplot(ktn, aes(eduyrs)) + geom_histogram(fill = "#A7AEDF", 
                 bins = 15)+labs(x = 'Education years')
#Descriptive Statistics
summary(ktn$eduyrs)
```

The years of full-time education completed variable means the total amount of time a person spent receiving an education regardless of its level. The number of years of full-time education completed has a normal distribution, with a mean of 11.68 years and a median of 11 years correspondingly.

#### Level of Religiosity

```{r warning = FALSE, message=FALSE}
ktn$rlgdgr <- as.numeric(as.character(ktn$rlgdgr)) 
ktn$rlgdgr_cat[ktn$rlgdgr =="0"|ktn$rlgdgr =="1"| ktn$rlgdgr =="2"| ktn$rlgdgr == "3"] <- "not religious" 
ktn$rlgdgr_cat[ktn$rlgdgr == "4" |ktn$rlgdgr =="5"| ktn$rlgdgr =="6" | ktn$rlgdgr =="7"] <- "mildly religous" 
ktn$rlgdgr_cat[ktn$rlgdgr >="8"] <- "very religious"

ktn$rlgdgr_cat_sorted <- factor(ktn$rlgdgr_cat, ordered = TRUE,
levels = c("not religious", "mildly religous", "very religious"))

#Plot
ggplot(na.omit(ktn), aes(x =rlgdgr_cat_sorted)) + geom_bar(fill = "#A7AEDF")+
  labs(x="Level of religiosity", y = "Number of people")
```

We decided to categorize the variable of religiosity. In the dataset there was the question "How religious are you?", so variable indicates subjective level of religiosity. Data included scale from 0 to 10, where 0 stands for "Not at all religious", and 10 is "Very religious". We assigned category "not religious" to answers from 0 to 3 (inclusive), "mildly religious" to answers from 4 to 6 (inclusive), and "very religious" to answers from 8 to 10 (inclusive).

#### Gender Answer Group

```{r}
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Female"] <- "Women about Women"
ktn$admge_cat[ktn$admge == "Group 1" & ktn$gndr == "Male"] <- "Men about Women"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Female"] <- "Women about Men"
ktn$admge_cat[ktn$admge == "Group 2" & ktn$gndr == "Male"] <- "Men about Men"

#Plot
ggplot(na.omit(ktn), aes(x =admge_cat)) + geom_bar(fill = "#A7AEDF")+
  labs(x="Gender Answer group", y = "Number of people")
```

By considering which answer group the person is in (group 1 or group 2) and comparing this to the gender of the person him/herself, we generated 4 subgroups, likewise evenly distributed, comprising about 380-400 people, with the category of women about men being the lowest with 357 people.

## Analysis

```{r warning = FALSE, message=FALSE}
view_df(ktn, show.string.values=TRUE)
```

## Correlation

### Statistical hypotheses:

-   H0: there is no monotonic relationship between ideal age to get married and ideal age to become a parent, ideal age to start living with a partner, years of full-time education, age of the respondent.

-   HA: there is monotonic relationship between ideal age to get married and ideal age to become a parent, ideal age to start living with a partner, years of full-time education, age of the respondent.

#### Ideal age to start living with a partner and Ideal age to get married

```{r warning = FALSE, message=FALSE}
ggscatter(ktn, x = "iagmr", y = "iagpnt", 
add = "reg.line", 
cor.coef = TRUE, 
cor.method = "pearson")

```

#### Ideal age to start living with a partner and Ideal age to get married

```{r warning = FALSE, message=FALSE}
ggscatter(ktn, x = "iagmr", y = "iaglptn", 
add = "reg.line", 
cor.coef = TRUE, 
cor.method = "pearson")

```

#### Education years and Ideal age to get married

```{r warning = FALSE, message=FALSE}
ggscatter(ktn, x = "iagmr", y = "eduyrs", 
add = "reg.line", 
cor.coef = TRUE, 
cor.method = "pearson")

```

#### Age and Ideal age to get married

```{r warning = FALSE, message=FALSE}
ggscatter(ktn, x = "iagmr", y = "age", 
add = "reg.line", 
cor.coef = TRUE, 
cor.method = "pearson")
```

From the plots above we can see that all relationships are linear.

### Normality assumption

```{r warning = FALSE, message=FALSE}
library("ggpubr")
par(mfrow=c(2,2))  

iagmr_plot = ggqqplot(ktn$iagmr, ylab = "iagmr",xlab="")
iagpnt_plot = ggqqplot(ktn$iagpnt, ylab = "iagpnt",xlab="")
iaglptn_plot = ggqqplot(ktn$iaglptn, ylab = "iaglptn",xlab="")
eduyrs_plot = ggqqplot(ktn$eduyrs, ylab = "eduyrs",xlab="")
age_plot = ggqqplot(ktn$age, ylab = "age",xlab="")
cowplot::plot_grid(iagmr_plot, 
                   iagpnt_plot, 
                   iaglptn_plot,
                   eduyrs_plot,
                   age_plot,
                   nrow = 3,
                   labels = "auto")

par(mfrow=c(3,2))  
hist(ktn$iagmr, xlab="Ideal age to get married")
hist(ktn$iagpnt, xlab="Ideal age to become a parent")
hist(ktn$iaglptn, xlab="Ideal age to start living with a partner")
hist(ktn$eduyrs, xlab="Education Years")
hist(ktn$age, xlab="Age")
 

```

From the normality plots and histograms, we conclude that all variables are not normally distributed, consequently we will further proceed with non-parametric coefficients, namely Spearman's rank correlation coefficient.

```{r warning = FALSE, message=FALSE}
cor.test(ktn$iagmr,  ktn$iagpnt, method = "spearman")

cor.test(ktn$iagmr,  ktn$iaglptn, method = "spearman")

cor.test(ktn$iagmr,  ktn$eduyrs, method = "spearman")

cor.test(ktn$iagmr,  ktn$age, method = "spearman")

```

### Correlation Matrix

```{r warning = FALSE, message=FALSE}
tab_corr(ktn[, c("iagmr", "iagpnt", "iaglptn", "eduyrs", "age")],
         corr.method = "spearman")
```

```{r warning = FALSE, message=FALSE}
sjp.corr(ktn[, c("iagmr", "iagpnt", "iaglptn", "eduyrs", "age")],
         corr.method = "spearman")
```

Based on the result of the test, we conclude that all correlations are statistically significant, low to strong, and indeed,monotonic relationships exists. To be more precise, we can conclude that:

-   Ideal age to get married and ideal age to become a parent are strongly positively correlated with a correlation coefficient of 0.772 and p-value of 2.2e-16 (\<0.05). Therefore, we reject H0 in favor of an alternative one.

-   Ideal age to get married and ideal age to start living with a partner are moderately positively correlated with a correlation coefficient of 0.523 and p-value of 2.2e-16 (\<0.05). Therefore, we reject H0 in favor of an alternative one.

-   Ideal age to get married and years of full-time education completed are weakly positively correlated with a correlation coefficient of 0.181 and p-value of 1.4e-11 (\<0.05). Therefore, we reject H0 in favor of an alternative one.

-   Ideal age to get married and age of a respondent are weakly negatively correlated with a correlation coefficient of -0.251 and p-value of 2.2e-16 (\<0.05). Therefore, we reject H0 in favor of an alternative one.

## Regression

### Plots of the variables

```{r warning = FALSE, message=FALSE}

cowplot::plot_grid(ggplot(ktn, aes(x = iagmr, y = age )) +
		                  geom_point() + 
		                  geom_smooth()+
                  labs(x = "Ideal age to get married", y = "Age"),
                  ggplot(ktn, aes(x = iagmr, y = eduyrs )) +
		                  geom_point() + 
	                  	geom_smooth()+
                  labs(x = "Ideal age to get married", y = "Years of full education"),
                  ggplot(ktn, aes(x = iagmr , y = hinctnta_cat )) +
		                  geom_point() + 
		                  geom_smooth()+
                  labs(x = "Ideal age to get married", y = "Income decile"),
                   nrow = 2,
                   labels = "auto")


cowplot::plot_grid(ggplot(na.omit(ktn), aes(x = eisced_cat_sorted, y = iagmr)) +
                      geom_boxplot() +
                      stat_summary(fun = "mean",colour = "#A7AEDF", size = 1)+
                  labs(x = "Level of education", y = "Ideal age for getting married"),
                  ggplot(na.omit(ktn), aes(x = rlgdgr_cat_sorted, y = iagmr)) +
                      geom_boxplot() +
                      stat_summary(fun = "mean",colour = "#A7AEDF", size = 1)+
                  labs(x = "Level of religiosity", y = "Ideal age for getting married"),
                  nrow = 1,
                   labels = "auto")


```

From the boxplots we observe that the mean and median values of ideal age for getting married are higher for higher education level in comparison to the lower and middle one. And regarding the religiosity, the mean and median values are decreasing as the religiosity level of the respondent is increasing.

### Model Comparison

Initially we would like to add all the control variables to our regression model, and then simplify, change and compare it till we reach the best model fit.

```{r warning = FALSE, message=FALSE}
s1 <- lm(iagmr ~ hinctnta_cat + age + eduyrs+eisced_cat + rlgdgr_cat+admge_cat, data = ktn) 
s2 <- lm(iagmr ~ hinctnta_cat + age +eisced_cat + rlgdgr_cat+admge_cat, data = ktn) 
tab_model(s1,s2,  show.ci = F)

```

As we observe, a better model fit is obtained with the removal of education years variables (11,6% compared to 11,5%). So, we will proceed with it. Now, we would also like to check how removal of Level of education variable will change the model.

```{r warning = FALSE, message=FALSE}
s3 <- lm(iagmr ~ hinctnta_cat + age +eisced_cat + rlgdgr_cat+admge_cat, data = ktn) 
s4 <- lm(iagmr ~ hinctnta_cat + age + rlgdgr_cat+admge_cat, data = ktn) 
tab_model(s3,s4,  show.ci = F)
```

We can conclude that Level of education removal doesn't really affect the model and only decreases the R2 a bit. Thus, we ended up keeping this variable.

### Model Interpretation

```{r warning = FALSE, message=FALSE}
s <- lm(iagmr ~ hinctnta_cat + age +eisced_cat + rlgdgr_cat+admge_cat, data = ktn) 
summary(s)
```

In our case, it can be seen that p-value of the F-statistic is \< 2.2e-16, which is highly significant. This means that, at least, one of the predictor variables is significantly related to the outcome variable.

The minimum residual was -10.5944, the median residual was -0.0813 and the max residual was 25.2254. Since the Min and Max, 1Q and 3Q are not matching (3rd quantile and Max values are larger), we can conclude that the distribution of residuals is more right-skewed.

In our model, the following variables were not statically significant:

-   the highest level of education (eisced_cat), p-value \> 0.05

-   level of religiosity (not religious category), p-value \> 0.05

As for the other variables, their coefficients were statically significant. More precisely:

1.  The **Income** of the respondents is statistically significant with p-value = 0.034 and estimated coefficient of 0.11. That way, we can see that the average increase in the indicated level of income is associated with an increase by 0.11 in the ideal age for getting married.
2.  The **Age** of the respondents is statistically significant with p-value = 3.97e-09 and estimated coefficient of -0.04. That way, we can see that the average increase in the age of the respondent is associated with an decrease by 0.04 in the ideal age for getting married.
3.  **Level of religiosity** is statistically significant with p-value = 0.021. This tells us that the ideal age for getting married is lower by 1.07 for respondents, who are very religious in comparison with those who mildly religious.
4.  **Gender answer group Men about Women** is statistically significant with p-value \< 0.001. This tells us that the ideal age for getting married is lower by 1.33 for male respondents answers about females in comparison with male respondents answers about men.
5.  **Gender answer group Women about Men** is statistically significant with p-value = 0.022 This tells us that the ideal age for getting married is higher by 0.79 for female respondents answers about males in comparison with male respondents answers about men.
6.  **Gender answer group Women about Women** is statistically significant with p-value = 0.018. This tells us that the ideal age for getting married is lower by 0.81 for female respondents who were answering about females in comparison with male respondents answering about men.

### Regression Equation

Finally, our model equation can be written as follow:

iagmr = 28.91 + 0.11 \* (hinctnta_cat) - 0.04 \* (age) - 1.07 \* (rlgdgr cat [very religious]) - 1.33 \* admge cat [Men about Women] + 0.79 \* (admge cat [Women about Men]) - 0.81 \* (admge cat [Women about Women])

```{r warning = FALSE, message=FALSE}
sjPlot::tab_model(s)
```

For our example, the R2 is 0.116, meaning that 11,6%(small proportion) of the variability of dependent variable (ideal age to get married) is being explained.

## Discussion

The correlation analysis confirms hypotheses number 3,5,6 and 7, and indeed we can conclude that the number of years spent on education, indicating the ideal age for becoming a parent and cohabitating, are positively correlated with the ideal age for getting married, as it was expected, since childbearing and partner living events are closely related to each other and people with higher levels of SES(considering education here as well) marry and start families later. As well, we observe a negative correlation of age and ideal marital age variables, which corresponds to the growing trend of desiring a later marriage.

The regression model analysis confirmed hypotheses number 1, 4, 7, 8, so we can conclude that the respondent's income, gender answer group, religiosity, and age are statistically significant, while the level of education has no significant effect on the perceived ideal age of marriage and thus the second hypothesis is not confirmed. To be more precise, as the level of religiosity and age of the respondent increases, the perceived ideal age to get married decreases, whereas greater income of the respondent is associated with a later ideal age of getting married.

Regarding gender answer subgroups, results confirm that women are generally ascribed a lower ideal marital age, as substantiated by evolutionary theory, and that women estimate higher ideal age of marriage for men, as they are concerned about their spouse being able to support the family, and men generally chose a lower ideal marriage age for women, being driven by ideas about their reproductive value and desire to date younger women.

# Bibliography

[1] Fuller, J. N., Frost, A. M. H., & Burr, B. K. (2015). Exploring the Impact of Religiosity and Socioeconomic Factors on Perceived Ideal Timing of Marriage in Young Adults. Journal of Student Research, 4(1), 120-129. <https://doi.org/10.47611/jsr.v4i1.208>

[2] Griskevicius, V., Delton, A. W., Robertson, T. E., & Tybur, J. M. (2011). Environmental contingency in life history strategies: The influence of mortality and socioeconomic status on reproductive timing. Journal of Personality and Social Psychology, 100(2), 241--254. <https://doi.org/10.1037/a0021082>

[3] Andrew Cherlin; Caitlin Cross-Barnet; Linda M. Burton; Raymond Garrett-Peters (2008). Promises They Can Keep: Low-Income Women's Attitudes Toward Motherhood, Marriage, and Divorce. , 70(4), 919--933. <doi:10.1111/j.1741-3737.2008.00536.x>

[4] Willoughby, B. J., & Carroll, J. S. (2011). Correlates of Attitudes Toward Cohabitation. Journal of Family Issues, 33(11), 1450--1476.doi:10.1177/0192513x11429666

[5] Arland Thornton; Linda Young-DeMarco (2001). Four Decades of Trends in Attitudes Toward Family Issues in the United States: The 1960s Through the 1990s. , 63(4), 1009--1037. <doi:10.1111/j.1741-3737.2001.01009.x>

[6] Todosijevic, Bojan, and Suzana Ignjatović. 2021. 'Gender Differences in Perception of the Appropriate Maturity Age for Men and Women: Age Norms of Reproduction-Related Life Events between the Social Context and Evolutionary Foundations'. Sociologija 63:289--313. doi: 10.2298/SOC2102289T.

